\documentclass[11pt]{report}
% \documentclass[12pt]{amsart}
% \documentclass[12pt]{amsproc}
% \documentclass[12pt]{amsbook}

\usepackage{amssymb,amscd,amsthm, verbatim,amsmath,color,fancyhdr, mathrsfs}
\usepackage{graphicx}
\usepackage{turnstile}
\usepackage[hidelinks]{hyperref}

\def\tequiv{\ensuremath\sdststile{}{}}

\usepackage[letterpaper, left=2cm, right=2cm, top=2.5cm,
bottom=2.5cm,dvips]{geometry}

\setcounter{section}{1}

\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{rem}[thm]{Remark}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{exr}[thm]{Exercise}
\newtheorem{ex}[thm]{Example}
\newtheorem{defn}[thm]{Definition}

\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\aB}{\mathcal{B}}
\newcommand{\aM}{\mathcal{M}}

\title{Measure and Probability theory}

\author{Victor Schmit}

\date{\today}

\begin{document}
\maketitle

\tableofcontents
\newpage

\chapter{Measures}

\input{leb}

\section{Stieljes Integral}
\label{sec:stieljes}

\section{Bounded Variation}

\defn[Total Variation]
Let $a : [0, T] \to \R$ be a function. For any $t \in [0,T]$, the \emph{total variation}
of $a$ on $[0,t]$ is defined by
\[
	V(t) =
	\sup \left\{
	\sum_{i=0}^{n}
	|a(t_{i+1}) - a(t_i)|
	\; : \;
	0 = t_0 < t_1 < \dots < t_{n+1} = t
	\right\}.
\]

\defn[Bounded Variation]
Let $a : [0, T] \to \R$ be a continuous function such that $a(0)=0$.
We say that $a$ is of \emph{bounded variation} on $[0,T]$ if
\[
	V(T) < +\infty .
\]

\thm
\label{thm:bound-stiel}
Let $a : [0,T] \to \R$ be continuous with $a(0)=0$.
Then $a$ is of bounded variation if and only if there exists a finite signed
measure $\mu$ on $[0,T]$ such that
\[
	a(t) = \mu([0,t]), \qquad t \in [0,T].
\]
In this case the total variation $V(T)$ coincides with the total variation of the
measure $\mu$.

\rem
This theorem is important because it allows us to interpret a function of bounded
variation as the cumulative mass of a signed measure. In other words, the
increments of the function correspond to the mass assigned by the measure:
\[
	a(t) - a(s) = \mu((s,t]).
\]

This provides a natural way to define integrals with respect to $a$.
If $f$ is a suitable function (for instance continuous), we define
\[
	\int_0^T f(t)\, da(t)
	:= \int_{[0,T]} f(t)\, \mu(dt).
\]

\rem
The integral with respect to a function of bounded variation can also be
defined as the limit of Riemann--Stieltjes sums.

Let
\[
	0 = t_0 < t_1 < \dots < t_n = T
\]
be a partition of $[0,T]$, and denote its mesh by
\[
	\max_{0 \le i \le n-1} (t_{i+1}-t_i).
\]

If $f$ is continuous and $a$ is of bounded variation, we define
\[
	\int_0^T f(t)\, da(t)
	=
	\lim_{\max (t_{i+1}-t_i) \to 0}
	\sum_{i=0}^{n-1}
	f(t_i)\big(a(t_{i+1})-a(t_i)\big).
\]

The bounded variation property ensures that the total size of the increments
\[
	\sum_{i=0}^{n-1} |a(t_{i+1})-a(t_i)|
\]
remains bounded. This prevents the Riemann sums from oscillating wildly and
ensures that the limit defining the integral exists.

To prove Theorem \ref{thm:bound-stiel}, we rely on several lemmas and
properties of the total variation function $V$.

\prop
Let $a : [0,T] \to \mathbb{R}$ and let $V$ denote its total variation
function.

\begin{enumerate}

	\item The function $V$ is non-decreasing and non-negative.
	\item For every $0 \le s < t \le T$,
	      \[
		      V(t) - V(s)
		      =
		      \sup \left\{
		      \sum_{i=0}^{n}
		      |a(t_{i+1}) - a(t_i)|
		      \ : \
		      s = t_0 < t_1 < \dots < t_{n+1} = t
		      \right\}.
	      \]

	      In other words, $V(t)-V(s)$ is the total variation of $a$ over the
	      interval $[s,t]$. We denote the total variation of $a$ over the interval
	      $[s,t]$ by $V_a(s,t)$.

	\item The function $V$ is continuous on $[0,T]$.
\end{enumerate}

\begin{proof}
	Let $a:[0,T]\to\R$ be a continuous function.

	\emph{(1)} We do not prove this point as it is very clear that $V$ is non-decreasing and non-negative.

	\medskip

	\emph{(2)} $V(t)-V(s)$ equals the variation of $a$ on $[s,t]$.
	Fix $0\le s < t\le T$ and define the variation of $a$ on $[s,t]$ by
	\[
		V_a(s,t)
		:=
		\sup\left\{
		\sum_{j=0}^{m-1}\bigl|a(x_{j+1})-a(x_j)\bigr|
		:\;
		s=x_0<\cdots<x_m=t
		\right\}.
	\]
	We prove that $V(t)-V(s)=V_a(s,t)$.

	\medskip

	\emph{Step 1: $V(t)-V(s)\ge V_a(s,t)$.}
	Let $\varepsilon>0$ and let $s=x_0<\cdots<x_m=t$ be an arbitrary partition of $[s,t]$.
	By definition of $V(s)$, there exists a partition $0=u_0<\cdots<u_n=s$ of $[0,s]$
	such that
	\[
		V(s)\le \sum_{i=0}^{n-1}\bigl|a(u_{i+1})-a(u_i)\bigr|+\varepsilon.
	\]
	Concatenate the two partitions to obtain a partition of $[0,t]$:
	\[
		0=u_0<\cdots<u_n=s=x_0<\cdots<x_m=t.
	\]
	Therefore, by definition of $V(t)$,
	\[
		V(t)\ge
		\sum_{i=0}^{n-1}\bigl|a(u_{i+1})-a(u_i)\bigr|
		+
		\sum_{j=0}^{m-1}\bigl|a(x_{j+1})-a(x_j)\bigr|.
	\]
	Subtracting the previous inequality for $V(s)$ yields
	\[
		V(t)-V(s)\ge
		\sum_{j=0}^{m-1}\bigl|a(x_{j+1})-a(x_j)\bigr|-\varepsilon.
	\]
	Since the partition of $[s,t]$ was arbitrary, taking the supremum over all such
	partitions gives
	\[
		V(t)-V(s)\ge V_a(s,t)-\varepsilon.
	\]
	Letting $\varepsilon\downarrow0$ we obtain $V(t)-V(s)\ge V_a(s,t)$.

	\medskip

	\emph{Step 2: $V(t)-V(s)\le V_a(s,t)$.}
	Let $\varepsilon>0$. By definition of $V(t)$, there exists a partition
	$0=t_0<\cdots<t_N=t$ of $[0,t]$ such that
	\[
		V(t)\le \sum_{k=0}^{N-1}\bigl|a(t_{k+1})-a(t_k)\bigr|+\varepsilon.
	\]
	Insert the point $s$ into this partition (if it is not already present), obtaining
	a refined partition $0=r_0<\cdots<r_M=t$ with $s=r_\ell$ for some $\ell$.
	Refinement can only increase the sum of increments, hence
	\[
		\sum_{k=0}^{N-1}\bigl|a(t_{k+1})-a(t_k)\bigr|
		\le
		\sum_{j=0}^{M-1}\bigl|a(r_{j+1})-a(r_j)\bigr|.
	\]
	Consequently,
	\[
		V(t)\le \sum_{j=0}^{M-1}\bigl|a(r_{j+1})-a(r_j)\bigr|+\varepsilon.
	\]
	Split the latter sum at $s=r_\ell$:
	\[
		\sum_{j=0}^{M-1}\bigl|a(r_{j+1})-a(r_j)\bigr|
		=
		\sum_{j=0}^{\ell-1}\bigl|a(r_{j+1})-a(r_j)\bigr|
		+
		\sum_{j=\ell}^{M-1}\bigl|a(r_{j+1})-a(r_j)\bigr|.
	\]
	The first sum is bounded by $V(s)$, and the second sum is bounded by $V_a(s,t)$.
	Therefore,
	\[
		V(t)\le V(s)+V_a(s,t)+\varepsilon,
	\]
	hence
	\[
		V(t)-V(s)\le V_a(s,t)+\varepsilon.
	\]
	Letting $\varepsilon\downarrow0$ gives $V(t)-V(s)\le V_a(s,t)$.
	Combining Steps 1 and 2 yields $V(t)-V(s)=V_a(s,t)$.

	\medskip

	\emph{(3)} $V$ is continuous. Let $0 \le s < t \le T$. Since $V$ is non–decreasing, the right limit
	\[
		V(s^+) := \lim_{u \downarrow s} V(u)
	\]
	and the left limit
	\[
		V(t^-) := \lim_{u \uparrow t} V(u)
	\]
	exist.
	Moreover,
	\[
		V(s) \le V(s^+) \le V(t^-) \le V(t).
	\]
	In particular,
	\[
		V(t^-) - V(s^+) \le V(t) - V(s).
	\]
	If we prove the reverse inequality, we obtain
	\[
		V(t)-V(s)=V(t^-)-V(s^+).
	\]
	Since $V(s)\le V(s^+)\le V(t^-)\le V(t)$, this would imply
	\[
		V(s^+)=V(s) \quad \text{and} \quad V(t^-)=V(t),
	\]
	hence $V$ would be continuous.
	We now prove the reverse inequality.
	Let $\varepsilon>0$ and let
	\[
		s=x_0 < x_1 < \dots < x_n=t
	\]
	be a partition of $[s,t]$.
	Since $a$ is continuous on $t$ and $s$, there exists $\delta>0$ such that
	\[
		|t-u|<\delta \quad \Longrightarrow \quad |a(t)-a(u)|<\varepsilon
		\quad \text{ and }\quad
		|s-v|<\delta \quad \Longrightarrow \quad |a(s)-a(v)|<\varepsilon .
	\]
	Refining the partition if necessary, we may assume that its mesh satisfies
	\[
		x_{k+1}-x_k<\delta .
	\]
	Now remove the endpoints $x_0=s$ and $x_n=t$ and denote
	\[
		y_0=x_1,\quad y_1=x_2,\quad \dots,\quad y_{n-2}=x_{n-1}.
	\]
	Then $(y_i)$ is a partition of $[y_0,y_{n-2}] \subset (s,t)$.
	Using the triangle inequality and the choice of $\delta$, we obtain
	\[
		\sum_{k=0}^{n-1}|a(x_{k+1})-a(x_k)|
		\le
		2\varepsilon + \sum_{k=0}^{n-3}|a(y_{k+1})-a(y_k)|.
	\]
	By the result of Point (2),
	\[
		\sum_{k=0}^{n-3}|a(y_{k+1})-a(y_k)|
		\le V(y_{n-2})-V(y_0).
	\]
	Since $y_0>s$ and $y_{n-2}<t$, we have
	\[
		V(y_{n-2})-V(y_0)\le V(t^-)-V(s^+).
	\]
	Therefore,
	\[
		\sum_{k=0}^{n-1}|a(x_{k+1})-a(x_k)|
		\le
		2\varepsilon + V(t^-)-V(s^+).
	\]
	Taking the supremum over all partitions $(x_i)$ of $[s,t]$ yields
	\[
		V(t)-V(s)\le 2\varepsilon + V(t^-)-V(s^+).
	\]
	Since $\varepsilon>0$ is arbitrary, we conclude that
	\[
		V(t)-V(s)\le V(t^-)-V(s^+).
	\]
	Combining this inequality with the reverse inequality obtained earlier gives
	\[
		V(t)-V(s)=V(t^-)-V(s^+).
	\]
	Hence $V$ is continuous on $[0,T]$.

\end{proof}

\rem
\label{rem:cont}
One can notice in the proof related to the continuity of the function $V$, that
$a$ is continuous to the right (to the left) implies that $V$ is continuous to the right (to the left).

\lem[Jordan decomposition]
\label{lem:jordan-cont}
Let $a:[0,T]\to\R$ be a continuous function of bounded variation with $a(0)=0$.
Then there exist two continuous non-decreasing functions $N,D:[0,T]\to\R$
such that
\[
	a(t)=N(t)-D(t), \qquad t\in[0,T],
\]
and
\[
	N(0)=D(0)=0.
\]

\begin{proof}
	Let $t \in [0,T]$. We define
	\[
		N(t) = \frac{V(t) + a(t)}{2}, \quad D(t) = \frac{V(t) - a(t)}{2}.
	\]
	Based on the definition of $V$, it is clear that $N,D$ are non-decreasing and non-negative.
	In addition, because $V$ and $a$ are continuous, $D$ and $N$ are also continuous. Finally,
	\[
		a(t)=N(t)-D(t)
	\]
	and
	\[
		N(0)=D(0)=0.
	\]
\end{proof}

\rem One can also notice that $V = N + D$.

\begin{proof}[Proof of Theorem \ref{thm:bound-stiel}]
	We want to prove that there exists a signed measure $\mu$ on $[0,T]$ such that $a(t) = \mu\left([0,t]\right)$.
	Based on the Jordan decomposition (Lemma \ref{lem:jordan-cont}), we have the following formula :
	\[
		a(t)=N(t)-D(t), \qquad t\in[0,T],
	\]
	with $N$ and $D$ continuous, non-decreasing, non-negative and that starts from $0$.
	Based on Section \ref{sec:stieljes} which deals with Stieljes integrals, there exists $\mu_1$ and $\mu_2$
	two positive measures on $[0,T]$ such that for every $t\in [0,T]$,
	\[
		N(t) = \mu_1([0,t]), \quad
		D(t) = \mu_2([0,t]).
	\]
	Finally, defining $\mu = \mu_1 - \mu_2$, a signed measure on $[0,T]$, we have
	\[
		a(t)=\mu([0,t]), \qquad t\in[0,T].
	\]
\end{proof}

\rem In fact, only right–continuity is required to represent $N$ and $D$ as cumulative
functions of positive measures. As noted in Remark \ref{rem:cont}, the total variation
function $V$ is right–continuous whenever $a$ is right–continuous. Consequently, the
Jordan decomposition still provides two non–decreasing right–continuous functions
$N$ and $D$, which can be represented as Stieltjes measures.

Therefore, the assumption that $a$ is continuous can be relaxed: it suffices to assume
that $a$ is right–continuous and of bounded variation. In this case $a$ admits a
representation as a Stieltjes integral, and in particular $a$ is a càdlàg function.

\end{document}
